Arc Blanc: a real time ocean simulation framework

### 2024-11-14 paper.tex
Precise the meaning

Let us introduce some notations as follows:
\begin{itemize}
	\item The scalar $H\in\mathbb R^+$ denotes the depth of the water. \Framework assumes \textbf{deep water} and therefore supposes that $H$ tends towards $+\infty$\footnote{See appendix \ref{sec:annexe} for more details on this assumption.}.
	\item The scalar $h(\mathbf x, t)\in\mathbb R$ is the height of the water surface at the horizontal position $\mathbf x=\left[x~z\right]^T$ and at time $t$.
	\item The vector $\mathbf{v}(\mathbf x, y, t) =  \left[v_x(\mathbf x, y, t)\ v_y(\mathbf x, y, t)\ v_z(\mathbf x, y, t)\right]^T\in\mathbb R^3$ represents the velocity of the fluid at the horizontal position $\mathbf x$, depth $y$ and time $t$.
	\item The scalar $\phi(\mathbf x,y,t)\in\mathbb R$ is the potential velocity of fluid at the horizontal position $\mathbf x$, depth $y$ and time $t$.
	\item The scalar $g=9.80665~m.s^{-2}$ denotes the standard gravity, while $\mathbf g\in\mathbb R^3$ is the gravitational field.
\end{itemize}

Tessendorf method is based on the following approximations of the Navier-Stokes equations\footnote{See appendix \ref{sec:annexe-NS}.}:
\begin{equation}
	\label{equa:systemTessendorf}
	\left\lbrace
	\begin{array}{ll}
		\frac{\partial\phi}{\partial t} = -gh(\mathbf x,t)            & \text{for }y=0,            \\
		\Delta\phi=0                                                  & \text{for }-H\leq y\leq 0, \\
		\frac{\partial h}{\partial t}=\frac{\partial\phi}{\partial y} & \text{for }y=0,            \\
		\frac{\partial\phi}{\partial y}=0                             & \text{for }y=-H.
	\end{array}
	\right.
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Water height}
\label{subsec:waterHeight}
Tessendorf (2004) \cite{tessendorfSimulatingOceanWater1999} proposed to solve Equation~\ref{equa:systemTessendorf} using a sum of waves of different amplitudes and wavelengths as follows:
\begin{equation}
	\label{equa:heightWater}
	h(\mathbf x,t)=\sum_{\mathbf{k}} \tilde{h}(t, \mathbf{k}) \exp\left( i \mathbf{k}\cdot \mathbf x\right),
\end{equation}
where $\mathbf{k}=\left[k_x\ k_z\right]^T$ is the wave vector with $k_x=\frac{2\pi n}{L_{i}}$ and $k_z=\frac{2\pi m}{L_{i}}$, $L_i$ is the length of the i-th cascade as defined in Section~\ref{subsec:cascades}, $n$ and $m$ are integers in $\left[-\frac{N}{2},\frac{N}{2}\right]$, $N\in\mathbb N^+$ being a constant given by user, and $\tilde{h}$ is given by:
\begin{equation}
	\label{equa:heightWaterBis}
	\tilde{h}(t, \mathbf{k}) = \tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)+\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t),
\end{equation}
where $k$ is the Euclidean norm of the wave vector $\mathbf{k}$, $\omega(k)=\sqrt{gk}$ is the \textbf{dispersion relation}\footnote{The dispersion relation is defined for deep water, see appendix \ref{sec:annexe} for more details on this assumption.} and $\tilde{h}_0^*$ the conjugate of $\tilde{h}_0$.
Researchers in oceanography have made statistical measurements on the ocean to compute $\tilde{h}_0$, leading to many \textit{ocean spectra} \cite{tessendorfGilliganPrototypeFramework2017}.
In its original paper, Tessendorf (2004) \cite{tessendorfSimulatingOceanWater1999} used the \textit{Phillips Spectrum} with Gaussian fluctuation, leading to the following expression of $\tilde{h}_0(\mathbf{k})$:
\begin{equation}
	\tilde{h}_0(\mathbf{k}) = \xi_{\mathbf{k}}\sqrt{\frac{4\pi}{L_ik}S(\omega)D(\omega, \theta)\left|\frac{\partial\omega(k)}{\partial k}\right|}
\end{equation}
where $\xi_{\mathbf{k}}$ are samples from a Gaussian distribution with mean $0$ and standard deviation $1$, $S(\omega)$ and $D(\omega, \theta)$ are the frequency and directional spectra described respectively in paragraphes \ref{subsubsec:jonswap} and \ref{subsubsec:donelanbanner}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Velocity of water}
\label{subsec:velWater}
Starting from the system \ref{equa:systemTessendorf}, \framework proposes to compute the velocity of the water at any depth $y$, as follows\footnote{See appendix \ref{sec:annexe} for more details on the origin of this expression}:
\begin{equation}\label{equa:velocity}
	% \left[
	%   \begin{array}{c} \displaystyle v_x(\mathbf x,y,t) \\ \displaystyle v_y(x,y,z,t) \\ \displaystyle v_z(x,y,z,t) \end{array}
	% \right]
	\mathbf v(\mathbf x,y,t)=
	\sum_{\mathbf{k}} \tilde{\mathbf{v}}(\mathbf{k},t,y)\exp\left(i\mathbf{k}\cdot\mathbf x\right),
\end{equation}
where the vector $\tilde{\mathbf{v}}(\mathbf{k},t, y)$ is given as follows:
\begin{equation}
	\tilde{\mathbf{v}}\left(\mathbf{k},t,y\right)=%
	E\left(\mathbf{k},y\right)\left(\tilde{h}_0\left(\mathbf{k}\right)\exp\left({i\omega(k)t}\right)-\tilde{h}_0^*\left(-\mathbf{k}\right)\exp\left({-i\omega(k)t}\right)\right)
	\left[
		\begin{array}{c}
			\displaystyle\frac{-k_xg}{\omega(k)} \\
			\displaystyle i\omega(k)             \\
			\displaystyle\frac{-k_zg}{\omega(k)}
		\end{array}
		\right],
\end{equation}
where the attenuation function $E(\mathbf{k},y)\in\mathbb R$ is defined as follows:
\begin{equation}
	\label{equa:attenuationDeep}
	E(\mathbf{k}, y)=
	\begin{cases}
		1+ky     & \text{if }y>0, \\
		\exp(ky) & \text{else}.
	\end{cases}
\end{equation}
The first case ($y>0$) comes from an extrapolation above the sea surface of the normal case ($y\leq0$).

>>>>

Let us introduce some notations as follows:
\begin{itemize}
	\item The scalar $H\in\mathbb R^+$ denotes the depth of the water. \Framework assumes \textbf{deep water} and therefore supposes that $H$ tends towards $+\infty$\footnote{See Appendix \ref{sec:annexe} for more details on this assumption.}.
	\item The scalar $h(\mathbf x, t)\in\mathbb R$ is the height of the water surface at the horizontal position $\mathbf x=\left[x~z\right]^T$ and at time $t$.
	\item The vector $\mathbf{v}(\mathbf x, y, t) =  \left[v_x(\mathbf x, y, t)\ v_y(\mathbf x, y, t)\ v_z(\mathbf x, y, t)\right]^T\in\mathbb R^3$ represents the velocity of the fluid at the horizontal position $\mathbf x$, depth $y$ and time $t$.
	\item The scalar $\phi(\mathbf x,y,t)\in\mathbb R$ is the potential velocity of fluid at the horizontal position $\mathbf x$, depth $y$ and time $t$.
	\item The scalar $g=9.80665~m.s^{-2}$ denotes the standard gravity, while $\mathbf g\in\mathbb R^3$ is the gravitational field.
\end{itemize}

Tessendorf method is based on the following approximations of the Navier-Stokes equations\footnote{See Appendix  \ref{sec:annexe-NS}.}:
\begin{equation}
	\label{equa:systemTessendorf}
	\left\lbrace
	\begin{array}{ll}
		\frac{\partial\phi}{\partial t} = -gh(\mathbf x,t)            & \text{for }y=0,            \\
		\Delta\phi=0                                                  & \text{for }-H\leq y\leq 0, \\
		\frac{\partial h}{\partial t}=\frac{\partial\phi}{\partial y} & \text{for }y=0,            \\
		\frac{\partial\phi}{\partial y}=0                             & \text{for }y=-H.
	\end{array}
	\right.
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Water height}
\label{subsec:waterHeight}
Tessendorf (2004) \cite{tessendorfSimulatingOceanWater1999} proposed to solve Equation~\ref{equa:systemTessendorf} using a sum of waves of different amplitudes and wavelengths as follows:
\begin{equation}
	\label{equa:heightWater}
	h(\mathbf x,t)=\sum_{\mathbf{k}} \tilde{h}(t, \mathbf{k}) \exp\left( i \mathbf{k}\cdot \mathbf x\right),
\end{equation}
where $\mathbf{k}=\left[k_x\ k_z\right]^T$ is the wave vector with $k_x=\frac{2\pi n}{L_{i}}$ and $k_z=\frac{2\pi m}{L_{i}}$, $L_i$ is the length of the i-th cascade as defined in Section~\ref{subsec:cascades}, $n$ and $m$ are integers in $\left[-\frac{N}{2},\frac{N}{2}\right]$, $N\in\mathbb N^+$ being a constant given by user, and $\tilde{h}$ is given by:
\begin{equation}
	\label{equa:heightWaterBis}
	\tilde{h}(t, \mathbf{k}) = \tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)+\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t),
\end{equation}
where $k$ is the Euclidean norm of the wave vector $\mathbf{k}$, $\omega(k)=\sqrt{gk}$ is the \textbf{dispersion relation}\footnote{The dispersion relation is defined for deep water, see Appendix~\ref{sec:annexe} for more details.} and $\tilde{h}_0^*$ the conjugate of $\tilde{h}_0$.
Researchers in oceanography have made statistical measurements on the ocean to compute $\tilde{h}_0$, leading to many different \textit{ocean spectra} \cite{tessendorfGilliganPrototypeFramework2017}.
In its original paper, Tessendorf (2004) \cite{tessendorfSimulatingOceanWater1999} used the \textit{Phillips Spectrum} with Gaussian fluctuation, leading to the following expression of $\tilde{h}_0(\mathbf{k})$:
\begin{equation}
	\tilde{h}_0(\mathbf{k}) = \xi_{\mathbf{k}}\sqrt{\frac{4\pi}{L_ik}S(\omega)D(\omega, \theta)\left|\frac{\partial\omega(k)}{\partial k}\right|}
\end{equation}
where $\xi_{\mathbf{k}}$ are samples from a Gaussian distribution with mean $0$ and standard deviation $1$, $S(\omega)$ and $D(\omega, \theta)$ are the frequency and directional spectra described respectively in paragraphes \ref{subsubsec:jonswap} and \ref{subsubsec:donelanbanner}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Velocity of water}
\label{subsec:velWater}
Starting from the system \ref{equa:systemTessendorf}, \framework proposes to compute the velocity of the water at any depth $y$, as follows\footnote{See Appendix   \ref{sec:annexe} for more details on the origin of this expression}:
\begin{equation}\label{equa:velocity}
	% \left[
	%   \begin{array}{c} \displaystyle v_x(\mathbf x,y,t) \\ \displaystyle v_y(x,y,z,t) \\ \displaystyle v_z(x,y,z,t) \end{array}
	% \right]
	\mathbf v(\mathbf x,y,t)=
	\sum_{\mathbf{k}} \tilde{\mathbf{v}}(\mathbf{k},t,y)\exp\left(i\mathbf{k}\cdot\mathbf x\right),
\end{equation}
where the vector $\tilde{\mathbf{v}}(\mathbf{k},t, y)$ is given as follows:
\begin{equation}
	\tilde{\mathbf{v}}\left(\mathbf{k},t,y\right)=%
	E\left(\mathbf{k},y\right)\left(\tilde{h}_0\left(\mathbf{k}\right)\exp\left({i\omega(k)t}\right)-\tilde{h}_0^*\left(-\mathbf{k}\right)\exp\left({-i\omega(k)t}\right)\right)
	\left[
		\begin{array}{c}
			\displaystyle\frac{-k_xg}{\omega(k)} \\
			\displaystyle i\omega(k)             \\
			\displaystyle\frac{-k_zg}{\omega(k)}
		\end{array}
		\right],
\end{equation}
where the attenuation function $E(\mathbf{k},y)\in\mathbb R$ is defined as follows:
\begin{equation}
	\label{equa:attenuationDeep}
	E(\mathbf{k}, y)=
	\begin{cases}
		1+ky     & \text{if }y>0, \\
		\exp(ky) & \text{else}.
	\end{cases}
\end{equation}
The first case ($y>0$) comes from an extrapolation above the sea surface of the normal case ($y\leq0$).
